Reparametrization of COM-Poisson Regression Models with Applications in the Analysis of Experimental Data

### 2017-09-18 article.Rnw
Improve simulation study section
 - Remove algorithm enviroment to show steps;
 - Include a figure to explain the choice of the parameters.

In this section we conducted a simulation study to assess the properties
of the proposed model and estimators0. We considered average counts varying
from 3 to 27 according to a regression model structure with a continuous
variable and a categorical variable with three levels. With respect to
the precision parameter $\phi$, we fixed it to four values $-1.6$,
$-1.0$, $0.0$ and $1.8$, in order to have strong overdispersed, moderate
overdispersed, equidispersed and underdispersed counts respectively. For
each value of precision parameter, we considered four different sample
sizes, 50, 100, 300 and 1000, generating 1000 datasets in each case. The
\autoref{alg:simulation} shows the details of the steps in simulation
study. In \autoref{fig:bias-plot} we show the bias of the estimators for
each scenario (combination between values of the dispersion parameter
and samples sizes) together with a confidence interval calculated as
average bias plus and minus $\Phi(0.975)$ times the standard error. The
scales are standardized for each parameter by dividing the bias by the
standard error obtained for the sample of size 50.

\begin{algorithm}
\caption{Steps in simulation study.}
\label{alg:simulation}
  \Begin{
  $\bm{\beta} = \begin{bmatrix} \beta_0 & \beta_1 & \beta_{21} &
    \beta_{22} \end{bmatrix}^\top = \begin{bmatrix} 2.0 &
    0.5 & 0.8 & -0.8 \end{bmatrix}^\top$\;
  \For{$n \in \{50, 100, 300, 1000\}$}{
    set $\bm{x}_1$ as a sequence, with $n$ elements, between $0$ and
    $1$\;
    set $\bm{x}_2$ as a repetition, with $n$ elements, of three
    categories\;
    compute $\bm{\mu}$ using $\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1
    +  \beta_{21} \bm{x}_{21} + \beta_{22} \bm{x}_{22})$, where
    $\bm{x}_{21}$ and $\bm{x}_{22}$ are \textit{dummy} variable for
    $\bm{x}_2$\;
    \For{$\phi \in \{-1.6, -1.0, 0.0, 1.8\}$}{
      \Repeat{$1000$ times}{
        simulate $\bm{y}$ from COM-Poisson distribution with $\bm{\mu}$
        and $\phi$ parameters\;
        fit COM-Poisson$_\mu$ regression model to $\bm{y}$ data with
        $\bm{X} = \begin{bmatrix} \bm{1} & \bm{x}_1 & \bm{x}_{21} &
          \bm{x}_{22} \end{bmatrix}$ design matrix\;
        get $\hat{\bm{\theta}} = \begin{bmatrix} \hat{\phi} & \hat{\beta}_0 &
          \hat{\beta}_1 & \hat{\beta}_{21} & \hat{\beta}_{22}
        \end{bmatrix}^\top$\;
        get confidence intervals for $\hat{\bm{\theta}}$ by quadratic
        approximation at the maximum likelihood estimate
        (assumes $\hat{\bm{\theta}} \sim \mathcal{N}(\hat{\bm{\theta}},
        \sqrt{-\bm{v}})$, where $\bm{v}$ is a diagonal of the inverse of
        the Hessian matrix)\;
      }
    }
  }
}
\end{algorithm}

@@

@@

The results in \autoref{fig:bias-plot} show that for all dispersed
levels, both the expected bias and standard error tend to 0 as the
sample size is increased. The estimators for the regression parameters
are unbiased, consistency and their empirical distributions are
symmetric. For the precision parameter, the estimator is asymptotically
unbiased, in small samples the parameter is overestimated and the
empirical distribution is right-skewed. \autoref{fig:coverage-plot}
presents the confidence interval coverage rate based on quadratic
approximation at the maximum likelihood estimate by sample size and
dispersed levels.

>>>>

In this section we conducted a simulation study to assess the properties
of the proposed model and estimators. We considered average counts
varying from 3 to 27 according to a regression model structure with a
continuous variable $\bm{x}_1$ between 0 and 1 and a categorical
variable $\bm{x}_2$ with three levels. Then the systematic componente is
given by $\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1 + \beta_{21}
\bm{x}_{21} + \beta_{22} \bm{x}_{22})$, where $\bm{x}_{21}$ and
$\bm{x}_{22}$ are \textit{dummy} variable for $\bm{x}_2$. With respect
to the precision parameter $\phi$, we fixed it to four values $-1.6$,
$-1.0$, $0.0$ and $1.8$, in order to have strong overdispersed, moderate
overdispersed, equidispersed and underdispersed counts
respectively. \autoref{fig:justpars} shows the variation of the average
counts (left) and variation of dispersion index (E$(Y)$ / var$(Y)$) for
each fixed value of the precision parameter $\phi$. This configuration
allows us to assess the proprierties of the estimators in extreme
situation such as This configuration allows us to evaluate the
properties of the estimators in extreme situations such as high counts
and low dispersion, and low counts and high dispersion.

In order to assess the consistency of the estimators we considered four
different sample sizes, 50, 100, 300 and 1000, generating 1000 datasets
in each case. In \autoref{fig:bias-plot} we show the bias of the
estimators for each scenario (combination between values of the
precision parameter and samples sizes) together with a confidence
interval calculated as average bias plus and minus $\Phi(0.975)$ times
the standard error. The scales are standardized for each parameter by
dividing the bias by the standard error obtained for the sample of size
50.

%-----------------------------------------------------------------------
% Explain the simulation
% \begin{algorithm}
% \caption{Steps in simulation study.}
% \label{alg:simulation}
%   \Begin{
%   $\bm{\beta} = \begin{bmatrix} \beta_0 & \beta_1 & \beta_{21} &
%     \beta_{22} \end{bmatrix}^\top = \begin{bmatrix} 2.0 &
%     0.5 & 0.8 & -0.8 \end{bmatrix}^\top$\;
%   \For{$n \in \{50, 100, 300, 1000\}$}{
%     set $\bm{x}_1$ as a sequence, with $n$ elements, between $0$ and
%     $1$\;
%     set $\bm{x}_2$ as a repetition, with $n$ elements, of three
%     categories\;
%     compute $\bm{\mu}$ using $\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1
%     +  \beta_{21} \bm{x}_{21} + \beta_{22} \bm{x}_{22})$, where
%     $\bm{x}_{21}$ and $\bm{x}_{22}$ are \textit{dummy} variable for
%     $\bm{x}_2$\;
%     \For{$\phi \in \{-1.6, -1.0, 0.0, 1.8\}$}{
%       \Repeat{$1000$ times}{
%         simulate $\bm{y}$ from COM-Poisson distribution with $\bm{\mu}$
%         and $\phi$ parameters\;
%         fit COM-Poisson$_\mu$ regression model to $\bm{y}$ data with
%         $\bm{X} = \begin{bmatrix} \bm{1} & \bm{x}_1 & \bm{x}_{21} &
%           \bm{x}_{22} \end{bmatrix}$ design matrix\;
%         get $\hat{\bm{\theta}} = \begin{bmatrix} \hat{\phi} & \hat{\beta}_0 &
%           \hat{\beta}_1 & \hat{\beta}_{21} & \hat{\beta}_{22}
%         \end{bmatrix}^\top$\;
%         get confidence intervals for $\hat{\bm{\theta}}$ by quadratic
%         approximation at the maximum likelihood estimate
%         (assumes $\hat{\bm{\theta}} \sim \mathcal{N}(\hat{\bm{\theta}},
%         \sqrt{-\bm{v}})$, where $\bm{v}$ is a diagonal of the inverse of
%         the Hessian matrix)\;
%       }
%     }
%   }
% }
% \end{algorithm}

@@

<<justpars, fig.height=4, fig.width=9, fig.cap="Average counts (left) and dispersion indexes (right) for each scenario considered in the simulation study.">>=

#-------------------------------------------
# Justify the choices of the parameters simulation
x1 <- seq(0, 1, length.out = 100)
x2 <- rep(letters[1:3], length.out = 100)
X <- model.matrix(~ x1 + x2)
mu <- exp(X %*% beta)

daexplain <- ldply(lapply(phis, function(phi) {
    va <- compute_variance(mu, phi, sumto = 300)
    data.frame(mu = mu, va = va, di = va / mu, x1 = x1, x2 = x2)
}), .id = "phi")

fl <- parse(text = gsub("=", "==", levels(daexplain$phi)))
xy1 <- xyplot(mu ~ x1, groups = x2,
              type = c("g", "l"),
              lwd = 2,
              xlab = "Continous variable",
              ylab = "Average count",
              auto.key = list(
                  column = 3,
                  points = FALSE,
                  lines = TRUE,
                  title = "Categorical variable",
                  cex.title = 1.1
              ))
xy2 <- xyplot(di ~ x1 | phi, groups = x2,
              type = c("g", "l"),
              lwd = 2,
              # scales = "free",
              layout = c(2, 2),
              xlab = "Continous variable",
              ylab = "Dispersion index",
              data = daexplain,
              strip = strip.custom(factor.levels = fl))

@@

The results in \autoref{fig:bias-plot} show that for all dispersed
levels, both the expected bias and standard error tend to 0 as the
sample size is increased. The estimators for the regression parameters
are unbiased, consistency and their empirical distributions are
symmetric. For the precision parameter, the estimator is asymptotically
unbiased, in small samples the parameter is overestimated and the
empirical distribution is slightly right-skewed.
\autoref{fig:coverage-plot} presents the confidence interval coverage
rate based on quadratic approximation at the maximum likelihood estimate
(assumes $\hat{\bm{\theta}} \sim \mathcal{N}(\hat{\bm{\theta}},
\sqrt{-\bm{v}})$, where $\bm{v}$ is a diagonal of the inverse of the
Hessian matrix) by sample size and dispersed levels.
